Boosting Higgs Pair Production in the $bb\bar{b}\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-06 paper/sec-analysis.tex
work on paper

@@

As explained above, in the boosted category we also require the two
leasing subjets of the large-$R$ jet to be relatively
hard.
%
In Fig ...

>>>>

\subsection{Kinematic distributions}

@@

As explained above, in the boosted category we also require the two
leasing subjets of the large-$R$ jet to be relatively
hard.
%
In Fig ... we show the distribution in $p_T$ of the leading
and subleading jets.
%
It is clear from the comparison that

Let's now consider the resolved category.
%
To begin with, here it is important to understand the $p_T$ distribution
of the four leading small-$R$ jets of the event.
%
Given that in general the boost from the Higgs decays is moderate,
the two subleading jets might be not too hard, and thus it
should be important to ensure that our $p_T^{\rm min}$ cut
is not too strong.
%
In Fig. we show ....

Also, it is useful to show the distribution in rapidity of the
individual jets

Finally, we show in
Fig.~\ref{fig:mhh} the invariant mass distribution of the Higgs pairs,
comparing the resolved category (left) with the boosted category (right)
%
This is interesting since it determines the degree of boost that the two
Higgs candidates will have after non-resonant production.
%
In the resolved case, we see that the distribution
in $m_{hh}$ is rather harder for the signal as for the background,
and thus one expects that cutting in $m_{hh}$ would help signal
discrimination: we will verify this with the MVA in the next section.
%
For the boosted category the trend of the $m_{hh}$ distribution
is very different because of the selection, with the
distribution now peaking at much higher values.
%
In this case signal and background distributions
look reasonably similar.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.48\textwidth]{plots/m_HH_res_C1.pdf}
  \includegraphics[width=0.48\textwidth]{plots/m_HH_boost_C1.pdf}
  \caption{\small Left plot: the invariant mass distribution of the Higgs
    pair candidates, $m_{hh}$, comparing signal and background events,
    in the resolved category.
    %
    Right plot: same for the boosted category.
}
\label{fig:mhh}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

### 2015-07-06 paper/sec-mcsamples.tex
work on paper

First of all in this section we discuss the details of the
Monte Carlo generation of the signal and background
events used in this work.

\subsection{Signal}
%%%%%%%%%%%%%%%%%%%%%%%%%%

Higgs pair production is simulated at leading order using
{\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca}, with finite $m_t$ effects taken into
account.\footnote{We thank Eleni Vrodynou for assistance here.}
%
We use the NNPDF 3.0 $n_f = 4$ LO set~\cite{Ball:2014uwa} with
$\alpha_S(m_Z^2)=0.118$
as interfaced via {\tt LHAPDF6}~\cite{Buckley:2014ana}.
%
We use the default settings of the renormalization and factorization
scale of the model.
%
The Higgs boson parameters are also the same in the default model,
in particular we use $m_H=125$ GeV, consistent with the latest
measurements from ATLAS and CMS.

%
Parton level events are then showered with the {\tt Pythia8} Monte
Carlo~\cite{Sjostrand:2007gs,Sjostrand:2014zea}, in particular with {\tt v8.201}.
%
We use the default settings in the parton shower, in particular
we use the Monash 2013 tune~\cite{Skands:2014pea}, based on the NNPDF2.3LO PDF set~\cite{Ball:2012cx}.
%
We have switched off hadronisation, since this allows a more transparent modeling of
$b$-tagging.
%
A complete simulation of $b$-tagging at the hadron level is difficult to achieve without a full
detector simulator, which is beyond the scope of this paper.
%
For a similar reason, we don't consider underlying event, multiple parton interactions and
pile-up, which can be taken of using a variety of subtraction methods~\cite{Cacciari:2009dp}.

The signal samples are rescaled so that the integrated cross-section matches that of the
NNLO calculation~\cite{deFlorian:2013jea}: this corresponds to using a $K$-factor $\sigma_{NNLO}/sigma_{LO}=2.3$.
%
Let us note that resummed NNLO+NNLL calculations for Higgs pair production have also become available recently~\cite{deFlorian:2015moa},
leading to a moderate enhancement as compared to the fixed-order NNLO calculation.

\subsection{Background}

Now we turn to discuss the Monte Carlo generation of the relevant background processes.
%
All background samples are generated with the {\tt SHERPA} event generator~\cite{Gleisberg:2008ta}, v2.1.1.
%
As in the case of the signal generation, the NNPDF 3.0 $n_f = 4$ LO set with strong coupling $\alpha_S=0.118$ is used for all samples.
Factorisation and renormalisation scales are set as $\mu_F=\mu_R=H_T/2$.

Samples!!!!

At the generation level the following basic cuts are applied.
%
Each final state particle in the hard process must have $p_T \ge 20$ GeV, and be located within $| \eta | \le 3.0$. All final state particles must be separated by a minimum $\Delta R_{\mathrm{min}} =0.1$.
%
We have checked that the generator-level cuts are loose enough as compared to the actual
analysis cuts.
%
Total cross-sections and details of the samples generated are shown in Table~\ref{tab:samples}.

%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Process &  Generator & $N_{\mathrm{evt}}$ & $\sigma_{\mathrm{tot}}$ \\
\hline
\hline
$pp \to HH$ &  {\tt MadGraph5\_aMC@NLO} & 100K & $1.729\times10^{-2}$ pb \\
\hline
\hline
$pp \to b\bar{b}b\bar{b}$ &  {\tt SHERPA}v2.1.1 & 3M &$1.121 \times10^3$ pb \\
$pp \to b\bar{b}jj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.659 \times 10^5$ pb \\
$pp \to jjjj$ &  {\tt SHERPA}v2.1.1 & 3M  & $9.709\times 10^6$ pb \\
$pp \to t\bar{t}$ &  {\tt SHERPA}v2.1.1 & 3M & $2.514\times 10^3$ pb \\
\hline
\end{tabular}
\caption{\small Summary of signal and background samples generated in this work,
together with the corresponding generator-level cross-section.} \label{tab:samples}
\end{center}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As a cross-check of the background cross-sections reported in Table~\ref{tab:samples}, we have produced multi-jet samples
using the {\tt MadGraph5\_aMC@NLO} program, and compared with the results for the same processes reported in
Ref.~\cite{Alwall:2014hca}.
%
For comparison with the latter numbers, 
we require in all samples four anti-$k_T$ $R=0.5$ jets with $p_T \ge 80 $ GeV, and the leading jet must have $p_T \ge 100$ GeV.
%
All jets must be within an acceptance of $|\eta| \le 2.5 $.
%
In the case of the samples with $b$ quarks in the final state, these requirements are extended to the appropriate number of $b$-jets. For example, in the 2$b$2$j$ sample there must be at least two $b$-jets that pass the cuts outlined above.
%
We find agreement, within the scale uncertainties, between the {\tt MadGraph5\_aMC@NLO} and {\tt SHERPA} calculations of the multi-jet
backgrounds, so we can be confident that the set-up that we will use in this analysis is robust enough.

>>>>

First of all in this section we discuss the settings of the
Monte Carlo generation of the signal and background
events used in this work.
%
We also discuss how we study the effects of finite detector
resolution by means of an smearing of the final-state
particles four-momentum.

\subsection{Higgs pair production}
%%%%%%%%%%%%%%%%%%%%%%%%%%

Higgs pair production is simulated at leading order using
{\tt MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca}, with finite $m_t$ effects taken into
account.\footnote{We thank Eleni Vrodynou for assistance here.}
%
We use the NNPDF 3.0 $n_f = 4$ LO set~\cite{Ball:2014uwa} with
$\alpha_S(m_Z^2)=0.118$
as interfaced via {\tt LHAPDF6}~\cite{Buckley:2014ana}.
%
We use the default settings of the renormalization and factorization
scale of the model.
%
The Higgs boson parameters are also the same in the default model,
in particular we use $m_H=125$ GeV, consistent with the latest
measurements from ATLAS and CMS~\cite{Aad:2014aba,Khachatryan:2014jba}.
%
The value of the Higgs trilinear coupling $\lambda$ is set to its
Standard Model value.
%
The calculation is performed in the
$N_f$=4 scheme and thus
takes into account the finite mass of the $b$-quarks.

The signal samples are rescaled so that the integrated cross-section matches that of the
NNLO calculation of the inclusive cross-section from Ref.~\cite{deFlorian:2013jea}:
this corresponds to using a $K$-factor $\sigma_{\rm NNLO}/\sigma_{\rm LO}=2.3$.
%
Let us note that resummed NNLO+NNLL calculations for Higgs pair production have also become available recently~\cite{deFlorian:2015moa},
leading to a moderate enhancement as compared to the fixed-order NNLO calculation.

%
Parton level events are then showered with the {\tt Pythia8} Monte
Carlo~\cite{Sjostrand:2007gs,Sjostrand:2014zea}, in particular with {\tt v8.201}.
%
We use the default settings in the parton shower, in particular
we use the Monash 2013 tune~\cite{Skands:2014pea}, based on the NNPDF2.3LO PDF set~\cite{Ball:2012cx}.
%
We have switched off hadronisation, since this allows a more transparent modeling of
$b$-tagging.
%
A complete simulation of $b$-tagging at the hadron level is difficult to achieve without a full
detector simulator, which is beyond the scope of this paper.
%
For a similar reason, we don't consider underlying event, multiple parton interactions and
pile-up, which can be taken of using a variety of subtraction methods~\cite{Cacciari:2009dp}.
%
The same applies to the background samples, which we discuss now

\subsection{Backgrounds: QCD multijet and $t\bar{t}$ production}

Now we turn to discuss the Monte Carlo generation of the relevant background processes.
%
All background samples are generated at leading-order
with the {\tt SHERPA} event generator~\cite{Gleisberg:2008ta}, v2.1.1.
%
As in the case of the signal generation, the NNPDF 3.0 $n_f = 4$ LO set with strong coupling $\alpha_S(m_Z^2)=0.118$ is used for all samples.
%
Factorisation and renormalisation scales are set as $\mu_F=\mu_R=H_T/2$ for all
the samples.

We have considered all the relevant samples that can lead to the event
being tagged as a $HH\to 4b$ candidate.
%
First of all QCD we have $4b$ multi-jet production of course.
%
But also we generate QCD $2b2j$ and $4j$ multi-jet samples,
that can lead to the event being tagged in the case of multiple light
jets being mistagged as $b$-quarks.
%
While the light jet mistag probability is small, we find that
in general the $2b2j$ and $4j$ backgrounds cannot be neglected because
of their large cross-sections and enhancement from combinatorics, that
increase their contribution as compared to a naive estimate.
%
Neglecting the  $2b2j$ and $4j$ backgrounds is particularly
important for the resolved and intermediate categories, but less
so for the boosted one.
%
We don't include the NLO $K$-factor for the backgrounds but instead assume
a given experimental overall systematic uncertainties common to all backgrounds.

In addition to the QCD multijet, we also generate $t\bar{t}$ samples
in the fully hadronic final state, which lead to a $2b4j$ signature that can
contribute to the fake rates, and that has a similar topology that
the corresponding QCD sample.
%
We have used a value of the top quark mass of $m_t=173.2$ GeV.
%
Semileptonic decays of top quarks can be easily removed by requiring
a lepton veto.

At the generation level the following basic cuts are applied to
background events.
%
Each final state particle in the hard process must have $p_T \ge 20$ GeV, and be located
in the central region with
$| \eta | \le 3.0$.
%
In addition, at the matrix element level
all final state particles must be separated by a minimum $\Delta R_{\mathrm{min}} =0.1$.
%
We have checked that the generator-level cuts are loose enough as compared to the actual
analysis cuts.
%

Total cross-sections and details of the samples generated are shown in Table~\ref{tab:samples}.
%
First of all we see that the $t\bar{t}$ and QCD $4b$ samples are of
the same order of magnitude.
%
The $bbjj$ cross-section is more than a factor 200 as compared to the
$4b$ result, so in principle it should be subleading taking into
account the mistag rate, but as we will show below,
this is not true due to combinatoric.
%
 For top quark production, only the hadronic final state is generated.
  %
The $hh$ cross-section includes the NNLO $K$-factor.

%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\hline
Process &  Generator & $N_{\mathrm{evt}}$ & $\sigma_{\mathrm{tot}}$ (pb) \\
\hline
\hline
$pp \to HH$ &  {\tt MadGraph5\_aMC@NLO} & 100K & $3.87\times10^{-2}$  \\
\hline
\hline
$pp \to b\bar{b}b\bar{b}$ &  {\tt SHERPA}v2.1.1 & 3M &$1.121 \times10^3$  \\
$pp \to b\bar{b}jj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.659 \times 10^5$  \\
$pp \to jjjj$ &  {\tt SHERPA}v2.1.1 & 3M  & $9.709\times 10^6$  \\
$pp \to t\bar{t}\to b\bar{b}jjjj$ &  {\tt SHERPA}v2.1.1 & 3M & $2.514\times 10^3$ \\
\hline
\end{tabular}
\caption{\small Summary of signal and background samples generated in this work,
  together with the corresponding generator-level cross-section.
  %
  For top quark production, only the hadronic final state is generated.
  %
The $hh$ cross-section includes the NNLO $K$-factor.  
} \label{tab:samples}
\end{center}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As a cross-check of the background cross-sections reported in Table~\ref{tab:samples}, we have produced leading order
multi-jet samples
using the {\tt MadGraph5\_aMC@NLO} program, and compared with the results for the same processes reported in
Ref.~\cite{Alwall:2014hca}.
%
For comparison with the latter numbers, 
we require in all samples four anti-$k_T$ $R=0.5$ jets with $p_T \ge 80 $ GeV, and the leading jet must have $p_T \ge 100$ GeV, and
also that all jets must be within an acceptance of $|\eta| \le 2.5 $.
%
We find agreement, within the scale uncertainties, between the {\tt MadGraph5\_aMC@NLO} and {\tt SHERPA} calculations of the multi-jet
backgrounds, so we can be confident that the set-up that we will use in this analysis is robust enough.

Finally, let us discuss the smearing applied to signal
and background samples.
%
It is beyond the scope of this work to perform a complete study of the impact of
detector effects in this process.
%
However, it is still important to include some form of rough estimate of detector
effects in the analysis, to properly estimate for example the impact
of the Higgs mass window cuts.
%
In this work we simulate the finite momentum resolution of the ATLAS and CMS
hadronic calorimeters by applying a Gaussian smearing to the four components
of the four-momentum $p$, with mean zero and standard deviation $\Delta p$.
%
We take as baseline value a smearing of $\Delta p=5\%$, and in
Sect.~\ref{sec:optimisation} we explore how our results change
if a more optimistic or pessimistic value is assumed for the momentum
resolution in the LHC detectors.
